---
title: "WW2 dataset Initial analysis description"
author: "Wojciech Artichowicz"
date: "22 kwietnia 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

library(ggmap)
library(maps)
library(mapdata)
```

## Initial analysis of the dataset files

#### Analysis pipeline
 - importing data
 - data cleaning
    * checking for duplicates
    * checking integrity of the data

### Importing weather stations datafile
```{r importing weater stations}
df_stations <- read.csv("data/Weather Station Locations.csv")
head(df_stations,3)
```

#### Dataset size:
```{r weater stations dataset size (num of records)}
nrow(df_stations)
```


#### Some weather stations are repeated, checking that
```{r weater stations dataset checking}
nt <- table(df_stations$NAME)
nt[nt>1]
df_stations[df_stations$NAME == "BASTIA/CORSICA" | df_stations$NAME == "SAN JOSE",]
```
ok it is not an error - stations do not duplicate


### Importing weather measurement datafile
```{r importing weather measurement datafile}
df_weather <-  read.csv("data/Summary of Weather.csv")
head(df_weather,3)
```

Set weather station as factor
#### overview weather measurement dataset plot
```{r setting STA as factor}
df_weather$STA <- as.factor(df_weather$STA)
```

```{r plotting weather measurement datafile}
ggplot(df_weather,aes(x = MinTemp, y = MaxTemp, col = STA)) + 
  geom_point() + scale_colour_discrete(guide = FALSE)
```
```{r dataset merging}
df_all <- merge(x = df_weather, y = df_stations, by.x = "STA", by.y = "WBAN")
```

```{r plotting weather measurement datafile with coutries}
ggplot(df_all,aes(x = MinTemp, y = MaxTemp, col = STATE.COUNTRY.ID)) + 
  geom_point() 
  
```

```{r removing rows for which MinTemp > MaxTemp}
 df_all <- df_all[df_all$MinTemp < df_all$MaxTemp,]

```


```{r plotting filtered data}
 ggplot(df_all,aes(x = MinTemp, y = MaxTemp, col = STATE.COUNTRY.ID)) + 
  geom_point() #+ geom_text(aes(col = STATE.COUNTRY.ID,label=STATE.COUNTRY.ID),hjust=0, vjust=0)
  
```

####Facet wrap of the max ~ min temperature 
```{r plotting filtered data on facet grid}
 ggplot(df_all,aes(x = MinTemp, y = MaxTemp, col = STATE.COUNTRY.ID)) +
  geom_point(show.legend = FALSE) + facet_wrap(~STATE.COUNTRY.ID) 
```
```{r plotting filtered data on facet grid free scales}
ggplot(df_all,aes(x = MinTemp, y = MaxTemp, col = STATE.COUNTRY.ID)) +
  geom_point(show.legend = FALSE) + facet_wrap(~STATE.COUNTRY.ID,scales="free")
```

```{r plotting filtered data on the worldmap}
wrld <- map_data("world")
ggplot() + geom_polygon(data = wrld, aes(x=long, y = lat, group = group)) +
  geom_point(data = df_stations, aes(x = Longitude, y = Latitude, col = STATE.COUNTRY.ID)) +   geom_text(data = df_stations, aes(x = Longitude, y = Latitude, col = STATE.COUNTRY.ID,label=STATE.COUNTRY.ID),hjust=0, vjust=0) + scale_fill_discrete(guide=FALSE) + scale_color_discrete(guide = FALSE)
  
```